### 2023-02-04 outputs/paper/paper.qmd
Analysis added

---
title: "Death License Registrations May Indicate a Discreptency in Reported Data During the COVID-19 Pandemic"
subtitle: "An Analysis of Death License Registrations in Toronto from 2016-2020"
author: "Chloe Thierstein"
thanks: "Code and data are available at: https://github.com/cthierst/death_registry_analysis.git."
date: "February 3, 2023"
date-format: "D MMMM YYYY"
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract:  "Death registry statistics offer us a useful tool to better out healthcare system, from monitoring trends to government policy, it is essential to our understanding of the world around us. The COVID-19 pandemic marked a significant shift in the Canadian mortality rate in the year 2020. This paper utilizes death registry data from three civic centres in the greater Toronto area from 2016 to 2020, alongside secondary sources and research. From this work, we demonstrate that even data from considerably reliable sources may become distorted if enough care is not taken in its recording and the world around the data is not considered."
bibliography: references.bib
fig-cap-location: top
---

# 1 Introduction

Canadian death registry statistics offer a useful tool in helping better our healthcare system as by monitoring trends in public health such as infectious diseases, suicide, and unintentional injuries they help the healthcare sector provide better services and resources, like screening and prevention programs [@citestatscan3]. However, maintaining this data may not always be done accurately which could cause confusion over our understanding of Canadian demography and health.

In this paper we look at death registry statistics from three civic centres in the greater Toronto area, Etobicoke, Scarborough and North York between 2016 and 2020. From this data, we find that the death licenses provided by these centres during the first year of the COVID-19 pandemic (2020), are much lower than in previous years [@citeopendatatoronto2]. This is significant as Canada experienced an increase of 7.7% in deaths in 2020 [@citestatscan4], and 5.2% more deaths than would typically be expected when taking into account Canada's aging population [@citestatscan5]. We reason that this discrepancy in expected data versus reported data could be the result of strain and complications put onto record-keeping during this time similarly to what was seen in early 2019 with long delays in the distribution of death, birth and marriage certificates due to high online demand [@citectvnews]. Future work could look specifically at how strain on data reporting centres during the COVID-19 pandemic influenced data and mitigation techniques for the future

This paper will begin with an overview of its data management, source and cleaning. Next we will briefly overview the drop off in trend from previous years in death registrations to begin our discussion. Next we will consider the relationship between death licenses provided at each civic centre over the period of 2016 to 2020, to better understand how the year 2020 differed from past years. Next, we discuss in more detail, death licenses provided by each civic centre to determine a possible cause for the uncharacteristic decline of death licenses in 2020. We will then look at the mean number of deaths from each civic centre to further gauge how these centres relate to rates of death registrations. Finally, we will consider the limitations of the data, including biases.

# 2 Data

## 2.1 Data Management

This paper utilizes the R statistical programming language [@citeR], along with several packages. These packages are, tidyverse [@citetidyverse], janitor [@citejanitor], here [@citehere] and dplyr [@citedplyr]. The data being analyzed comes from Open Data Toronto and it is imported using the opendatatoronto package [@citeopendatatoronto]. All figures have been created using ggplot2 [@citeggplot2] and the tables have been created with knitr [@citeknitr] and kableExtra [@citekablextra], packages. The color styles in the graphs were created by using the RColorBrewer [@citercolorbrewer] package and any graph combinations were made using the Patchwork [@citepatchwork] package.

## 2.2 Data Source and Cleaning

The data comes from death registrations which are entered into the Registry Services Tracking System (RSTS) by Registry Services staff who are located at three of the civic centres, Etobicoke, North York and Scarborough [@citeopendatatoronto2]. It's creation supports the Vital Statistics Act, a Provincial legislation [@citeopendatatoronto2] which involve the collection of deaths, marriages, stillbirths, and live births [@citestatscan3]. This data set is updated monthly [@citeopendatatoronto2]. The variables from this data set represent the civic centres; "Etobicoke", "North York", and "Scarborough", number of death licenses registered in the month, place of death; "Outside City Limits" or "Toronto", and time period by month and year in which the death was registered [@citeopendatatoronto2].

To properly analyze this data for the purposes of this paper certain data was removed including, all data for the years 2011-2015 and 2021-2023. This was done to ensure that the data set being worked with had sufficient and consistent data to draw from without being too broad in scope. Additionally, rows from the civic centre variable labelled "Toronto" were removed as death registrations at this location were especially low and its data input, sporadic. Finally, the variable for place of death was not included in the analysis of this data as it was not relevant to its narrative.

## 2.3 Data Analysis

```{r}
#| include: false

#### Workspace Set-Up ####

#### Load Library ####
library(tidyverse)
library(opendatatoronto)
library(janitor)
library(ggplot2)
library(dplyr)
library(knitr)
library(RColorBrewer)
library(kableExtra)
library(patchwork)
library(readr)
library(here)

#### Import cleaned_death_registry.csv ####
cleaned_death_registry_data <- read.csv(here::here("./inputs/data/cleaned_death_registry.csv"))

#### Turning Years into Factors ####
cleaned_death_registry_data$Year <- as.factor(cleaned_death_registry_data$Year)

```

The province of Ontario represents a large portion of deaths in Canada, taking on an estimated 37% of Canadian mortality from 2016-2020 [@citestatscan1]. It is also estimated that on average, that just over 100 thousand people have died each year in Ontario between 2016 and 2020 [@citestatscan1]. We may suggest then, that Toronto makes up a large portion of this mortality rate, as it makes up approximately 44% of the Ontarian population [@citestatscan2]. Additionally, Canada has an aging population, meaning that it's mortality rate is expected to rise every year although it rose exponentially due to the COVID-19 pandemic [@citestatscan4]. Therefore, we may suspect that the COVID-19 pandemic would increase the mortality rate significantly in Toronto as Canada saw an increase of 7.7% in deaths in 2020 [@citestatscan4].

```{r}
#| echo: false
#| warning: false
#| label: fig-1
#| fig-cap: "Death Licenses Registered at Three Toronto Civic Centres from 2016-2020"

#### Creating Graph to Compare Years Against Death Licenses Given Out From 2016-2020 ####

ggplot(data=cleaned_death_registry_data, aes(x = Year, y = death_licenses, fill = Year)) +
  geom_bar(stat = "identity") +
  theme_classic() + 
  labs(
    x = "Year",
    y = "# of Death Licenses"
  ) +
  theme(legend.position = "none")+
  scale_fill_brewer(palette="Set2")

```

When looking at @fig-1, however, we can see that from 2016 to 2018 to there was an increase in death licenses registered, as expected, followed by a slight drop in 2019. This drop in 2019 may be indicative of a backlog from Ontario's Vital Statistics in early 2019 which was cited as having delays of over three months, waiting for death, birth and marriage certificates [@citectvnews]. What is most important in @fig-1 is that when looking at the "2020" bar in @fig-1, we can see that the death licenses registered drops significantly in comparison to any of the previous 4 months. However, the amount of death licenses registered each month in 2020 was inconsistent as can be seen in @fig-2.

```{r}
#| echo: false
#| warning: false
#| label: fig-2
#| fig-cap: "Death Licenses Registered by Month at Three Toronto Civic Centres from 2016-2020"

#### Creating Graph to Compare Months Against Death Licenses Given Out From 2016-2020, Including Visualization of Years ####

#### Ordering Months #####
cleaned_death_registry_data$Month <- factor(cleaned_death_registry_data$Month, levels=c("Jan", "Feb","Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

ggplot(data=cleaned_death_registry_data, aes(x = Month, y = death_licenses, fill = Year)) +
  theme_classic() +
  geom_bar(stat = "identity") +
  labs(
    x = "Month",
    y = "# of Death Licenses"
  ) +
  scale_fill_brewer(palette = "Set2")
```

\newpage

By looking at @fig-2 we can see that, generally, each month is relatively consistent in the amount of death licenses registered from the three civic centres from 2016 to 2019 with a few variations. However, when looking at 2020, we notice significantly lower rates of death license registration each month overall in comparison to past years with the exception of May and November. In the later, we see death license registrations that are nearly balanced with previous years. Whereas in May, there is a large spike in death license registration that far surpasses any month's death license registration in the previous four years.

Looking at May's spike without knowing the context of the years, we may assume that it is a particularly deadly month, however that we can see that the spike in May, shown in @fig-2, is likely indicative of the first wave of COVID-19 in Canada hitting its peak at the beginning of the month before declining [@citejamanetwork]. This makes sense, however, COVID-19 infections were not the only factor driving the increased death rates in Canada during 2020. There were indirect consequences of the pandemic which lead to the higher death rate in Canada. For example, there were over four-thousand deaths from accidental poisonings which was nearly one-thousand more than in 2019 [@citestatscan4]. By understanding that the increase in the Canadian death rate in 2020 was not only caused by COVID-19 infections, we can see the complexity of the situation. Although the continued decline in death licenses registered from these civic centres during a year that saw an unexpected 5.2% [@citestatscan5] rise in deaths indicates a potential issue in the reporting done at these civic centres.

@fig-3 demonstrates how many death licenses were registered at each civic centre, Etobicoke, North York and Scarborough from 2016 to 2020, both monthly and yearly. From @fig-3 we can see that, generally, Scarborough registers the most death licenses, followed by Etobicoke and finally North York. Interestingly, in 2020, Etobicoke largely overtook Scarborough in death license registrations while simulataneously registered more death licenses than it had in the previous four years. This could indicate that the Etobicoke centre was taking on more work from other civic centres as it was perhaps less overwhelmed during the COVID-19 pandemic. However, more data would be needed to confirm this.

```{r}
#| include: false
#| message: false

#### Creating Graph to Compare Months Against Death Licenses Given Out From 2016-2020 From Each Civic Centre ####

month_by_death_license <- ggplot(data=cleaned_death_registry_data, aes(x = Month, y = death_licenses, fill = civic_centre)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(stat = "identity") +
  theme_classic() + 
  labs(
    x = "Month",
    y = "# Death Licenses",
    fill = "Civic Centre"
  ) +
  scale_fill_brewer(palette="Set2")+
  theme(legend.position = "none") + #include legend at bottom to ensure readability when using patchwork package
  theme(axis.title.y = element_text(size = 10)) +
  theme(axis.title.x = element_text(size = 10))

```

```{r}
#| include: false
#| message: false

#### Creating Graph to Compare Years Against Death Licenses Given Out From 2016-2020 From Each Civic Centre ####

year_by_death_license <-  ggplot(data=cleaned_death_registry_data, aes(x = Year, y = death_licenses, fill = civic_centre)) + #assign ggplot to new variable in order to use patchwork package later
  geom_bar(stat = "identity") +
  theme_classic() + 
  labs(
    x = "Year",
    y = "# Death Licenses",
    fill = "Civic Centre"
  )+
  scale_fill_brewer(palette="Set2")+
  theme(legend.position = "bottom") + #include legend at bottom to ensure readability when using patchwork package
  theme(axis.title.y = element_text(size = 10))+
  theme(axis.title.x = element_text(size = 10))
```

```{r}
#| echo: false
#| warning: false
#| label: fig-3
#| fig-cap: "Monthly and Yearly Number of Death Licenses Registered by Each Civic Centre from 2016-2020"

#### Using Patchwork Package to Combine Two Graphs ####

(month_by_death_license) / (year_by_death_license) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 1)
```

```{r}
#| include: false

#### Creating Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####

#### Repeat Process for Each Year (2016-2020) ####

#### 2016 ####
et_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
  filter(Year == "2016" | civic_centre == "Etobicoke") |> #filter for year and civic centre
  group_by(Month) |> #group by month
  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out

#### 2017 ####
et_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
  filter(Year == "2017" | civic_centre == "Etobicoke") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2018 ####
et_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
  filter(Year == "2018" | civic_centre == "Etobicoke") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2019 ####
et_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
  filter(Year == "2019" | civic_centre == "Etobicoke") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2020 ####
et_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
  filter(Year == "2020" | civic_centre == "Etobicoke") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

```

```{r}
#| include: false

#### Combining Datasets for 5 Year Period by Month at Etobicoke Civic Centre ####

#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####

#### 2016 ####
et_all_avg_num_deaths_by_month_data <- #new dataset
  et_avg_num_deaths_by_month_2016 |> #start with first dataset in this spot
  left_join(et_avg_num_deaths_by_month_2016, by = "Month") #use left_join() to combine dataset

#### 2017 ####
et_all_avg_num_deaths_by_month_data <- 
  et_all_avg_num_deaths_by_month_data |> #for second and all following datasets use "new" dataset in this position
  left_join(et_avg_num_deaths_by_month_2017, by = "Month")

#### 2018 ####
et_all_avg_num_deaths_by_month_data <- 
  et_all_avg_num_deaths_by_month_data |>
  left_join(et_avg_num_deaths_by_month_2018, by = "Month")

#### 2019 ####
et_all_avg_num_deaths_by_month_data <- 
  et_all_avg_num_deaths_by_month_data |>
  left_join(et_avg_num_deaths_by_month_2019, by = "Month")

#### 2020 ####
et_all_avg_num_deaths_by_month_data <-
  et_all_avg_num_deaths_by_month_data |>
  left_join(et_avg_num_deaths_by_month_2020, by = "Month")

```

```{r}
#| warning: false
#| echo: false
#| label: tbl-1
#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Etobicoke Centre"

#### Creating a Table to Demonstrate Average Number of Deaths Per Month from 2016-2020 at Etobicoke Centre ####

et_all_avg_num_deaths_by_month_data |>
  select(
    Month,
    "Average # of Deaths Registered.y",
    "Average # of Deaths Registered.x.x",
    "Average # of Deaths Registered.y.y",
    "Average # of Deaths Registered.x.x.x",
    "Average # of Deaths Registered.y.y.y"
  ) |> #previous step created a duplicate column, use select() to remove the second column in dataset
  kable(
    col.names = c(#rename columns to be more meaningful
      "Month",
      "2016",
      "2017",
      "2018",
      "2019",
      "2020"),
    digits = 0,
    booktabs = TRUE,
    linesep = ""
  ) |>
  add_header_above(c(" " = 1, "# of Death License Registrations" = 5))

```

```{r}
#| include: false

#### Creating Datasets for 5 Year Period by Month at Scarborough Civic Centre ####

#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####

#### 2016 ####
sc_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
  filter(Year == "2016" | civic_centre == "Scarborough") |> #filter for year and civic centre
  group_by(Month) |>  #group by month
  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out

#### 2017 ####
sc_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
  filter(Year == "2017" | civic_centre == "Scarborough") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2018 ####
sc_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
  filter(Year == "2018" | civic_centre == "Scarborough") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2019 ####
sc_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
  filter(Year == "2019" | civic_centre == "Scarborough") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2020 ####
sc_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
  filter(Year == "2020" | civic_centre == "Scarborough") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

```

```{r}
#| include: false

#### Combining Datasets for 5 Year Period by Month at Scarborough Civic Centre ####

#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####

#### 2016 ####
sc_all_avg_num_deaths_by_month_data <- #new dataset
  sc_avg_num_deaths_by_month_2016 |> #start with first dataset in this spot
  left_join(sc_avg_num_deaths_by_month_2016, by = "Month") #use left_join() to combine dataset

#### 2017 ####
sc_all_avg_num_deaths_by_month_data <- 
  sc_all_avg_num_deaths_by_month_data |> #for second and all following datasets use "new" dataset in this position
  left_join(sc_avg_num_deaths_by_month_2017, by = "Month")

#### 2018 ####
sc_all_avg_num_deaths_by_month_data <- 
  sc_all_avg_num_deaths_by_month_data |>
  left_join(sc_avg_num_deaths_by_month_2018, by = "Month")

#### 2019 ####
sc_all_avg_num_deaths_by_month_data <- 
  sc_all_avg_num_deaths_by_month_data |>
  left_join(sc_avg_num_deaths_by_month_2019, by = "Month")

#### 2020 ####
sc_all_avg_num_deaths_by_month_data <-
  sc_all_avg_num_deaths_by_month_data |>
  left_join(sc_avg_num_deaths_by_month_2020, by = "Month")

```

```{r}
#| echo: false
#| warning: false
#| label: tbl-2
#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at Scarborough Centre"

#### Creating a Table to Demonstrate Average Number of Deaths Per Month From 2016-2020 at Scarborough Centre ####

sc_all_avg_num_deaths_by_month_data|>
  select(Month, "Average # of Deaths Registered.y","Average # of Deaths Registered.x.x", "Average # of Deaths Registered.y.y", "Average # of Deaths Registered.x.x.x", "Average # of Deaths Registered.y.y.y")|> #previous step created a duplicate column, use select() to remove the second column in dataset
  kable(
    col.names = c( #rename columns to be more meaningful
      "Month",
      "2016",
      "2017",
      "2018",
      "2019",
      "2020"
    ),
    digits = 0,
    booktabs = TRUE,
    linesep = "") |>
  add_header_above(c(" " = 1, "# of Death License Registrations" = 5))

```

```{r}
#| include: false

#### Creating Datasets for 5 Year Period by Month at North York Civic Centre ####

#### Repeat Process for Each Year (2016-2020) ####

#### 2016 ####
ny_avg_num_deaths_by_month_2016 <-cleaned_death_registry_data |>
  filter(Year == "2016" | civic_centre == "North York") |> #filter for year and civic centre
  group_by(Month) |> #group by month
  summarize("Average # of Deaths Registered" = mean(death_licenses)) #find average of death licenses given out

#### 2017 ####
ny_avg_num_deaths_by_month_2017 <-cleaned_death_registry_data |>
  filter(Year == "2017" | civic_centre == "North York") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2018 ####
ny_avg_num_deaths_by_month_2018 <-cleaned_death_registry_data |>
  filter(Year == "2018" | civic_centre == "North York") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2019 ####
ny_avg_num_deaths_by_month_2019 <-cleaned_death_registry_data |>
  filter(Year == "2019" | civic_centre == "North York") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

#### 2020 ####
ny_avg_num_deaths_by_month_2020 <-cleaned_death_registry_data|>
  filter(Year == "2020" | civic_centre == "North York") |>
  group_by(Month) |> 
  summarize("Average # of Deaths Registered" = mean(death_licenses))

```

```{r}
#| include: false

#### Combining Datasets for 5 Year Period by Month at North York Civic Centre ####

#### After 1st Instance (2016), Repeat 2nd Instance Process for Each Year (2017-2020) ####

#### 2016 ####
ny_all_avg_num_deaths_by_month_data <- #new dataset
  ny_avg_num_deaths_by_month_2016 |> #start with first dataset in this spot
  left_join(ny_avg_num_deaths_by_month_2016, by = "Month") #use left_join() to combine dataset

#### 2017 ####
ny_all_avg_num_deaths_by_month_data <- 
  ny_all_avg_num_deaths_by_month_data |>#for second and all following datasets use "new" dataset in this position
  left_join(ny_avg_num_deaths_by_month_2017, by = "Month")

#### 2018 ####
ny_all_avg_num_deaths_by_month_data <- 
  ny_all_avg_num_deaths_by_month_data |>
  left_join(ny_avg_num_deaths_by_month_2018, by = "Month")

#### 2019 ####
ny_all_avg_num_deaths_by_month_data <- 
  ny_all_avg_num_deaths_by_month_data |>
  left_join(ny_avg_num_deaths_by_month_2019, by = "Month")

#### 2020 ####
ny_all_avg_num_deaths_by_month_data <-
  ny_all_avg_num_deaths_by_month_data |>
  left_join(ny_avg_num_deaths_by_month_2020, by = "Month") 
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-3
#| tbl-cap: "Average Number of Deaths Per Month from 2016-2020 at North York Centre"

#### Creating a Table to Demonstrate Average Number of Deaths Per Month From 2016-2020 at North York Centre ####

ny_all_avg_num_deaths_by_month_data|>
  select(Month, "Average # of Deaths Registered.y","Average # of Deaths Registered.x.x", "Average # of Deaths Registered.y.y", "Average # of Deaths Registered.x.x.x", "Average # of Deaths Registered.y.y.y")|> #previous step created a duplicate column, use select() to remove the second column in dataset
  kable(
    col.names = c( #rename columns to be more meaningful
      "Month",
      "2016",
      "2017",
      "2018",
      "2019",
      "2020"
    ),
    digits = 0,
    booktabs = TRUE,
    linesep = "") |>
  add_header_above(c(" " = 1, "# of Death License Registrations" = 5)) 
```

\newpage

Many of the trends spoken about previously can be seen explicitly in @tbl-1, @tbl-2, and @tbl-3. These tables present the average number of death licenses registered per month from 2016 to 2020 at each civic centre. These tables allow us to better grasp the variability in death licenses registered each month from 2016-2020. High variations in amount of death licenses registered can be seen mostly between months, although between years these variations become much smaller. This is important as upon first glance one would not assume inherently that the data actually shows much less death license registrations in 2020.

While more data and research needs to be done to fully understand the story of data collection and reporting during the COVID-19 pandemic, this paper looks at a screenshot of death registry data from three civic centres between 2016 and 2020. By doing this, we attempt to better understanding the significance of data collection and how it may be impacted by outside conditions. We suggest that there is data missing from the report due to a variety of factors and reason that pressure and complications but onto data-keeping centres during the COVID-19 pandemic may be to blame. We encourage future inquiries to look specifically at how strain put on data reporting centres during the COVID-19 pandemic influenced data and how to mitigate these strains in the future.

\clearpage

# References